Inferring the gravitational potential of the Milky Way with a few precisely measured stars

### 2014-03-05 text/paper2/ms.tex
working on method section

\subsection{Scalings}

In \paperone, we define a set of coordinates relative to the instantaneous phase-space position of the progenitor, $(\bs{r}_{p},\bs{v}_{p})$, normalized by the instantaneous tidal radius and escape velocity. Here we (1) instead scale the velocity coordinates by the velocity dispersion of the satellite, (2) reduce the dimensionality by considering only the magnitude of these vectors, and (3) take the logarithm of the magnitude,
\begin{equation}
  \ln R = \ln\left(\frac{\left\vert\bs{r} - \bs{r}_{p}\right\vert}{R_{\rm tide}}\right)\,\,\,\,,\,\,\,\,
  \ln V = \ln\left(\frac{\left\vert\bs{v} - \bs{v}_{p}\right\vert}{\sigma_v}\right).
\end{equation}
Figure~\ref{fig:R_V} shows distributions of $\ln R$ and $\ln V$ --- computed at $t=t_{\rm ub}$ for each particle --- as a function of progenitor mass from the simulations described in Section~\ref{sec:sims}. In these scaled coordinates we model distributions in $R$ and $V$ as log-normal and ignore covariances between $R$ and $V$. We find that for masses $\lesssim10^9M_\odot$, the variance ($\sigma^2$ in plot panels) remains approximately constant and the mean stays fixed at 0. For the largest mass simulation ($2.5\times10^9M_\odot$), the perturbation to the total potential due to the progenitor halo becomes significant and thus a large fraction of the stars are disrupted through interactions with the subhalo.

\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{../../plots/paper2/R_V.pdf}
\caption{ TODO }\label{fig:R_V}
\end{center}
\end{figure}

The scale-radius and velocity dispersion scale with the mass of the progenitor. [...observational errors] [ $D_{\rm ps}$]
[Cold vs. Hot streams in the context of observational errors]

@@

For all tests in this article, we model the potential of the parent galaxy as a three-component sum of a Miyamoto-Nagai disk \citep{}, Hernquist bulge \citep[spheroid;][]{}, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
\begin{align}
	&\Phi_{\rm disk}(R,z) = -\frac{GM_{\rm disk}}{\sqrt{R^2 + (a + \sqrt{z^2 + b^2})^2}}\\
	&\Phi_{\rm spher}(r) = -\frac{GM_{\rm spher}}{r + c}\\
	&\Phi_{halo}(x,y,z) = v_{h}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
\end{align}
where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ axis
ratios ($q_1$, $q_2$) and orientation of the halo with respect to the
baryonic disk ($\phi$):
\begin{align}
  C_1 &= \frac{\cos^2\phi}{q_1^2} + \frac{\sin^2\phi}{q_2^2}\\
  C_2 &= \frac{\sin^2\phi}{q_1^2} + \frac{\cos^2\phi}{q_2^2}\\
  C_3 &= 2\sin\phi\cos\phi \left(q_1^{-2} - q_2^{-2}\right).
\end{align}

>>>>

\subsection{Tidal debris}

For a satellite galaxy on a mildly eccentric orbit, stars are disrupted via tidal ``pruning'' due to the tidal field of the parent system or tidal shocking due to the rapidly changing tidal field as the progenitor moves through pericenter. The interplay between these two processes as a function of orbital properties will be studied in forthcoming work, and thus we focus on the slow disruption of stars due to steady tidal forcing.

Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a point-mass of mass $M\gg m$ \citep[the restricted three-body problem, e.g.,][]{binneytremaine}. In a frame rotating with the orbital frequency of the satellite, the effective potential around the satellite has two stationary, unstable points located around $\sim R_p \pm r_J$, where $R_p$ is the distance to the satellite and $r_J$ is the Jacobi or tidal radius,
\begin{equation}
	r_J = \left(\frac{m}{3M}\right)^{1/3}R_p.
\end{equation}
Particles that would be bound to the satellite system in isolation may have enough (Jacobi) energy to overcome the effective potential barrier at these Lagrange points and thus will be preferentially stripped from the satellite at these points (Fig.~8.6 of BT).

For a spherical, extended mass distribution [...] \label{eq:tidalradius}

[Describe velocity dispersion...] \label{eq:velscale}

In a triaxial potential (e.g., Equation~\ref{eq:lm10}), the orbital plane of the satellite is not fixed but we still expect there to be instantaneous Lagrange points along the line of centers connecting the origin of the parent potential to the satellite. We thus proceed by using coordinates relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane, and normalized by the instantaneous tidal radius for positions (Eq.~\ref{eq:tidal radius}) and instantaneous velocity scale for velocity (Eq.~\ref{eq:velscale}). 
% where $\hat{\bs r}_p = {\bs r}_p / |{\bs r}_p|$ and $\hat{\bs v}_p = {\bs v}_p / |{\bs v}_p|$
The (time-dependent) basis vectors are given by
\begin{align}
	\hat{x}_1 &= \bs{r}_p / \|\bs{r}_p\|\\
	\hat{x}_2 &= -\hat{x}_1 \times \hat{x}_3\\
	\hat{x}_3 &= (\bs{r}_p \times \bs{v}_p) / \|\bs{r}_p \times \bs{v}_p\|.
\end{align}
Figure~\ref{fig:lpts_r} shows sections of particle orbits in this coordinate system, symmetric around the unbinding time, $t_{\rm ub}$, (defined above), as a function of progenitor mass for each N-body simulation described in Section~\ref{sec:}. In these coordinates, the classical Lagrange points would be located around $\pm1$. Figure~\ref{fig:lpts_v} shows the same for the velocity of each star. The location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored. Furthermore, the dispersion in velocity is well-normalized by [...]. Even in [...], the phase-space dispersion of tidal debris scales with the mass ratio $(m_{\rm sat} / M_{\rm enc})^{1/3}$.

\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{../../plots/paper2/Lpts_r.png}
\caption{ TODO }\label{fig:lpts_r}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{../../plots/paper2/Lpts_v.png}
\caption{ TODO }\label{fig:lpts_v}
\end{center}
\end{figure}

[Cold vs. Hot streams in the context of observational errors]

@@

For all tests in this article, we model the potential of the parent galaxy as a three-component sum of a Miyamoto-Nagai disk \citep{}, Hernquist bulge \citep[spheroid;][]{}, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
\begin{align}
	&\Phi_{\rm disk}(R,z) = -\frac{GM_{\rm disk}}{\sqrt{R^2 + (a + \sqrt{z^2 + b^2})^2}}\\
	&\Phi_{\rm spher}(r) = -\frac{GM_{\rm spher}}{r + c}\\
	&\Phi_{\rm halo}(x,y,z) = v_{h}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
\end{align}
where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ axis
ratios ($q_1$, $q_2$) and orientation of the halo with respect to the
baryonic disk ($\phi$):
\begin{align}
  C_1 &= \frac{\cos^2\phi}{q_1^2} + \frac{\sin^2\phi}{q_2^2}\\
  C_2 &= \frac{\sin^2\phi}{q_1^2} + \frac{\cos^2\phi}{q_2^2}\\
  C_3 &= 2\sin\phi\cos\phi \left(q_1^{-2} - q_2^{-2}\right).
\end{align}
The total potential then is just
\begin{equation}
	\Phi_{\rm tot} = \Phi_{\rm disk} + \Phi_{\rm spher} + \Phi_{\rm halo}\label{eq:lm10}.
\end{equation}
